Optimizing Structural Properties Of Neural Networks With Genetic Algorithms

### 2015-08-01 paper/paper.tex
completed some text

@@

\section{Introduction}
In contemporary artificial neuronal network science a vast number of
different models for describing both the dynamic of networks as well as
their plasticities are applied. 
But what exactly makes artificial neuronal networks learn successfully? And
why are certain types of networks and certain learning rules effective for
some problems, but fail at solving others? Confronting and eventually
answering these questions is critical when trying to gain a deep
understanding of neural networks and artificial intelligence.

Using the above questions as a guideline,
we decided to analyze the FORCE algorithm introduced in \cite{FORCE} that
is especially suitable for learning periodic patterns. More precisely, we
wanted to identify which properties of a network are important for its
learning success, and how to choose these properties' values for an optimal
result.  To this end, one option would be to take an initial network and
an adequate periodic pattern to be learned, and to optimize the network's
internals for this very task. Then patterns may emerge that hint at how an
optimal network design looks like.

@@

\section{Concepts and Methods}
\paragraph{Basic Network Dynamics} As our neuronal network model we choose
basic rate networks that are also used in the original FORCE publication
\cite{FORCE} and the publication \cite{FORCE}, where our notation of this
paragraph mainly stems from. So we look at

- describe network dynamic
- describe learning and tasks
- describe sucess of network

- We chose rate networks in order to apply FORCE rule
- borrow notation from reward modulation rule paper
Basic Network dynamics:
    - Introduce single neuronal network as collection of matrices with specified
    dynamics
    - Briefly describe the learning rule (FORCE)

\paragraph{Genes and Challenges}
    - Introduce parameterized random networks as generators of these networks
    - Introduce challenges, and the expected success values when paring
    generators with challenges.

@@

\paragraph{Tasks}
For the optimization to work properly, the problems presented to the networks 
have (for the most part) to solvable by FORCE learning. While it was shown \cite{}
that large recurrent neural networks are capable of learning very complex functions
using the FORCE algorithm, the optimization requires a large number of learnings.
Therefore, the networks have to be restricted to being quite small ($\approx 100$ neurons)
and as such are less powerful.

@@

\begin{thebibliography}{}

\bibitem{mcdermott}
  Y.-F.~Chen {\it et al.},
  %``Microwave Photon Counter Based on Josephson Junctions,''
  Phys.\ Rev.\ Lett.\  {\bf 107}, 217401 (2011).

\end{thebibliography}

>>>>

    \newcommand{\ix}[1]{_\mathrm{#1}}
    \newcommand{\Ix}[1]{^\mathrm{#1}}

@@

\section{Introduction}
Looking at the field of artificial neuronal networks, one finds a great
variety of different models describing both the dynamic of networks as well
as their plasticities, i.e.\ their learning behavior.
But what exactly makes artificial neuronal networks learn successfully? And
why are certain types of networks and certain learning rules effective for
some problems, but fail at solving others? Confronting and eventually
answering these questions is critical when trying to gain a deep
understanding of neural networks and artificial intelligence.

Using the above questions as a guideline, we decided to analyze what
network characteristics makes the FORCE algorithm, introduced in
\citep{FORCE}, especially suitable for learning periodic patterns.  More
precisely, we wanted to identify which properties of a network are
important for its learning success, and how to choose these properties'
values for an optimal result.  To this end, one option would be to take an
initial network and an adequate periodic pattern to be learned, and to
optimize the network's internals for this very task. Then patterns may
emerge that hint at how an optimal network design looks like.

@@

\section{Concepts and Methods}
\paragraph{Network Model} We chose the same rate based network architecture
that is used in the original FORCE publication \citep{FORCE} (denoted as
architecture A) and also in \citep{RM}, where our notation mainly stems
from. So we look at a network $\mathcal{N}$ with $N$ internal neurons and
states $x_i$ for $1 \le i \le N$ that loosely represent the neurons
membrane potential. The firing rate $r_i$ of the $i$-th internal neuron is
given by $r_i = \tanh x_i$.  Further, the network contains $R$ readout
neurons with states $z_i$ for $1 \le i \le R$, where
\begin{equation} 
    z_i = \sum_{j=1}^{N}\omega\Ix{read}_{ij}\,r_j~.
\end{equation}
Taking into account both internal dynamics as well as an external feedback
pathway, the dynamics $t \mapsto x_i(t)$ of the single neurons are governed
by \footnote{~Note that we did not include external input signals in this
formula, as is done in \citep{FORCE} and \citep{RM}. We in fact enabled
inputs in our code base, but did not use them for the main results.}
\begin{equation}
    \dot{x}_i(t) = -x_i(t) + \sum_{i=1}^N \omega\Ix{rec}_{ij}\, r_j(t) + 
                   \sum_{i=1}^R \omega\Ix{fb}_{ij}\, z_j(t)~,
\end{equation}
where we introduce the internal recurrent synapse weights
$\omega\Ix{rec}_{ij}$ and the feedback synapse weights
$\omega\Ix{fb}_{ij}$. One can see that the neurons exponentially
decay if they are not presented stimulus by other neurons. 
We solved this system of differential equations by
a simple Newton method with integration time steps $dt = 0.01$, which was
also used in \citep{FORCE}. (How we constructed $\omega\Ix{rec}$ and
$\omega\Ix{fb}$ will be discussed in \emph{Parametric Random Networks})

\paragraph{Learning} Learning took place by stepwisely adapting the weights
$\omega\Ix{read}$ so that the dynamics $t\mapsto r_i(t)$ of the readout
rates should eventually match a given periodic input pattern $t\mapsto
a_i(t)$ as accurately as possible. To update $\omega\Ix{read}$ we chose
a variant of the FORCE algorithm \citep{FORCE} that rapidly modifies the
feedback loop to keep the errors $|r_i(t) - a_i(t)|$ small from the
beginning on.

While the authors of \citep{FORCE} update the weights $\omega\Ix{read}$ in
fixed intervals (every second time step), we used an adaptive mechanism to
determine how often learning should occur. This was initially introduced
for performance reasons, as the single learning steps proved to be the main
computational bottleneck, but we also found that the long time network
dynamics tended to be slightly more accurate for adaptive learning steps.

\paragraph{Parametric Random Networks -- Genotypes} Since we are not
interested in single network properties but rather want to find out which
structural elements of networks are important for learning success, we
introduced \emph{parametric random networks}, or \emph{network genotypes}.
Such a genotype $G$ consists of a set $\Theta$ of parameters that determine
which concrete networks are created when $G$ is expressed in a single
phenotype. Technically $G$ may be understood as a stochastic network
generator that returns a network $\mathcal{N}=G(\omega)$ for every seed
value $\omega$.
The genotypes' parameters $\theta\in\Theta$ were either integer or floating
point values and can be classified as affecting either (1) the topology of
$\omega\Ix{rec}$ or (2) the weight values of $\omega\Ix{rec}$ and
$\omega\Ix{fb}$. Important examples for (1) are the networks size $N$, the
occupation probability $p$ (for random Erdös-Renyi networks), the
neighborhood range $k$ (for ring topologies), or the number $l$ of layers
(for feed-forward networks). We furthermore introduced ratio parameters
that allowed us to interpolate between different topologies.  Parameters of
type (2) were the feedback strength \texttt{feedback} (scaling of
$\omega\Ix{fb}$) and the gain value \texttt{gain} (scaling of
$\omega\Ix{rec}$ by ).

\paragraph{Fitness} For a given genotype $G$ and a given challenge $C$, the
fitness $f_C(G)$ of $G$ for $C$ is taken to be the expected value of
the success rates of $G$'s phenotypes when solving the tasks of the
challenge $C$. In our simulations, $f_C(G)$ was of course only
approximated; this was usually done by taking 20 to 100 sample phenotypes
and sample tasks, where the exact number depended on the standard deviation
of the previously drawn samples' success rates.

@@

\paragraph{Tasks and Challenges}
For the optimization to work properly, the problems presented to the networks 
have (for the most part) to solvable by FORCE learning. While it was shown \citep{}
that large recurrent neural networks are capable of learning very complex functions
using the FORCE algorithm, the optimization requires a large number of learnings.
Therefore, the networks have to be restricted to being quite small ($\approx 100$ neurons)
and as such are less powerful.

@@

\bibliography{sources}
